---
title: "Cytoscape and igraph"
author: "Alex Pico"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cytoscape and igraph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette will show you how to convert networks between igraph and Cytoscape.

### Installation
```{r, echo=F, eval=F}
source("https://bioconductor.org/biocLite.R")
biocLite("RCy3", "igraph")
```

### Required Software
The whole point of RCy3 is to connect with Cytoscape. You will need to install and launch Cytoscape: 

* [Download Cytoscape](http://www.cytoscape.org/download.php) 
* Complete installation wizard
* Launch Cytoscape 

### From igraph to Cytoscape
The igraph package is a popular network tool among R users. With RCy3, you can easily translate igraph networks to Cytoscape networks!

Here is a basic igraph network construction from the [graph_from_data_frame docs](http://igraph.org/r/doc/graph_from_data_frame.html):
```{r,eval=F}
actors <- data.frame(name=c("Alice", "Bob", "Cecil", "David",
                            "Esmeralda"),
                     age=c(48,33,45,34,21),
                     gender=c("F","M","F","M","F"))
relations <- data.frame(from=c("Bob", "Cecil", "Cecil", "David",
                               "David", "Esmeralda"),
                        to=c("Alice", "Bob", "Alice", "Alice", "Bob", "Alice"),
                        same.dept=c(FALSE,FALSE,TRUE,FALSE,FALSE,TRUE),
                        friendship=c(4,5,5,2,1,1), advice=c(4,5,5,4,2,3))
g <- graph_from_data_frame(relations, directed=TRUE, vertices=actors)
```

You now have an igraph network, g.  In order to translate the network together with all vertex (node) and edge attributes over to Cytoscape, simply use:

```{r, eval=F}
createNetworkFromIgraph(g, "myNetwork")
```

### From Cytoscape to igraph 
Inversely, you can use createIgraphFromNetwork() in RCy3 to retreive vertex (node) and edge data.frames to construct an igraph network. 

```{r, eval=F}
g2 <- createIgraphFromNetwork("myNetwork")
```
