---
title: "Cytoscape and iGraph"
author: "by Alexander Pico"
package: RCy3
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    toc_float: true
vignette: >
  %\VignetteIndexEntry{2. Cytoscape and iGraph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
  eval=FALSE
)
```

This vignette will show you how to convert networks between iGraph and Cytoscape.

### Installation
source("https://bioconductor.org/biocLite.R")
biocLite("RCy3")
```{r}
library(igraph)
```

### Required Software
The whole point of RCy3 is to connect with Cytoscape. You will need to install and launch Cytoscape: 

* Download the latest Cytoscape from http://www.cytoscape.org/download.php 
* Complete installation wizard
* Launch Cytoscape 

```{r}
cytoscapePing()
```

### From igraph to Cytoscape
The igraph package is a popular network tool among R users. With RCy3, you can easily translate igraph networks to Cytoscape networks!

Here is a basic igraph network construction from the graph_from_data_frame docs, http://igraph.org/r/doc/graph_from_data_frame.html:
```{r}
actors <- data.frame(name=c("Alice", "Bob", "Cecil", "David",
                            "Esmeralda"),
                     age=c(48,33,45,34,21),
                     gender=c("F","M","F","M","F"))
relations <- data.frame(from=c("Bob", "Cecil", "Cecil", "David",
                               "David", "Esmeralda"),
                        to=c("Alice", "Bob", "Alice", "Alice", "Bob", "Alice"),
                        same.dept=c(FALSE,FALSE,TRUE,FALSE,FALSE,TRUE),
                        friendship=c(4,5,5,2,1,1), advice=c(4,5,5,4,2,3))
ig <- graph_from_data_frame(relations, directed=TRUE, vertices=actors)

# if function not found, then you need to install igraph. Try library(igraph)
```

You now have an igraph network, g.  In order to translate the network together with all vertex (node) and edge attributes over to Cytoscape, simply use:

```{r}
createNetworkFromIgraph(ig,"myIgraph")
```

### From Cytoscape to igraph 
Inversely, you can use createIgraphFromNetwork() in RCy3 to retreive vertex (node) and edge data.frames to construct an igraph network. 

```{r}
ig2 <- createIgraphFromNetwork("myIgraph")
```

Compare the round-trip result for yourself...
```{r, eval=F}
ig
ig2
```
