---
title: "Cytoscape and graphNEL"
author: "Alex Pico"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cytoscape and graphNEL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette will show you how to convert networks between graphNEL and Cytoscape.

### Installation
```{r, echo=F, eval=F}
source("https://bioconductor.org/biocLite.R")
biocLite("RCy3", "graph")
```

### Required Software
The whole point of RCy3 is to connect with Cytoscape. You will need to install and launch Cytoscape: 

* [Download Cytoscape](http://www.cytoscape.org/download.php) 
* Complete installation wizard
* Launch Cytoscape 

### From graphNEL to Cytoscape
The graph package is a popular network tool among R users. With RCy3, you can easily translate graphNEL networks to Cytoscape networks!

Create, display and render the 3-node, 3-edge simple graph
```{r cars}
demoNetworkFromGraph = function (base.url=.defaultBaseUrl) {
    window.title = 'demo.simpleGraph'
    if (window.title %in% as.character(getNetworkList (base.url)))
        deleteNetwork(window.title, base.url)
    
    g.simple = makeSimpleGraph()
    net.suid = createNetworkFromGraph(g.simple, window.title, base.url = base.url)
    layoutNetwork('force-directed', network=net.suid, base.url = base.url)
    setNodeLabelMapping ('label', network=net.suid, base.url = base.url)
    
    node.attribute.values = c("kinase",  "transcription factor")
    colors =                c('#A0AA00', '#FF0000')
    setNodeBorderWidthDefault(5, network=net.suid, base.url = base.url)
    setNodeBorderColorMapping(cws, 'type', node.attribute.values, colors, 
                              mode='lookup', default.color='#88FF22', 
                              network=net.suid, base.url = base.url)
    count.control.points = c(2, 30, 100)
    sizes                = c(20, 50, 100)
    setNodeSizeMapping('count', count.control.points, sizes, mode='interpolate',
                       network=net.suid, base.url = base.url)
    setNodeColorMapping('lfc', c(-3.0, 0.0, 3.0), 
                        c('#00FF00', '#FFFFFF', '#FF0000'), 
                        mode='interpolate',
                        network=net.suid, base.url = base.url)
    return(net.suid)
} # demoNetworkFromGraph
```

