---
title: "Cytoscape and graphNEL"
author: "Alex Pico"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cytoscape and graphNEL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette will show you how to convert networks between graphNEL and Cytoscape.

### Installation
```{r, echo=F, eval=F}
source("https://bioconductor.org/biocLite.R")
biocLite("RCy3", "graph")
```

### Required Software
The whole point of RCy3 is to connect with Cytoscape. You will need to install and launch Cytoscape: 

* [Download Cytoscape](http://www.cytoscape.org/download.php) 
* Complete installation wizard
* Launch Cytoscape 

### From graphNEL to Cytoscape
The graph package is a popular network tool among R users. With RCy3, you can easily translate graphNEL networks to Cytoscape networks!

Create a simple GraphNEL object
```{r}
g <- makeSimpleGraph()
```

Now pass it along to Cytoscape:

```{r, eval=F}
createNetworkFromGraph(g,"myGraph")
```
### From Cytoscape to GraphNEL 
Inversely, you can use createGraphFromNetwork() in RCy3 to retreive vertex (node) and edge data.frames to construct a GraphNEL object. 

```{r, eval=F}
g2 <- createGraphFromNetwork("myGraph")
```

Compare the round-trip result for yourself...
```{r, eval=F}
g
g2
```
