---
title: "Introduction to RCy3"
author: "by Alexander Pico"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to RCy3}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, echo = FALSE}
knitr::opts_chunk$set(
#  highlight = FALSE,
  eval=FALSE
)
```

This vignette will show you how to connect to Cytoscape from R using CyREST.  It will cover the installation of required packages and demonstrate some basic functionality of CyREST, commands and RCy3

It is recommended that your run this vignette *prior* to Cytoscape Automation workshops to mitigate troubleshooting time.

### Installation
source("https://bioconductor.org/biocLite.R")
biocLite("RCy3")

### Required Software
The whole point of RCy3 is to connect with Cytoscape. You will need to install and launch Cytoscape: 

* Download the latest Cytoscape from http://www.cytoscape.org/download.php
* Complete installation wizard
* Launch Cytoscape 

### First Contact
These functions are a convenient way to verify a connection to Cytoscape and for logging the versions of CyREST and Cytoscape in your scripts.

```{r}
cytoscapePing()
cytoscapeVersionInfo()
```

### My First Network
Let's create a Cytoscape network from some basic R objects
```{r}
nodes <- data.frame(id=c("node 0","node 1","node 2","node 3"),
           group=c("A","A","B","B"), # categorical strings
           score=as.integer(c(20,10,15,5)), # integers
           stringsAsFactors=FALSE)
edges <- data.frame(source=c("node 0","node 0","node 0","node 2"),
           target=c("node 1","node 2","node 3","node 3"),
           interaction=c("inhibits","interacts","activates","interacts"),  # optional
           weight=c(5.1,3.0,5.2,9.9), # numeric
           stringsAsFactors=FALSE)

createNetworkFromDataFrames(nodes,edges)
```

### Switch Styles
Check out the marquee style!
```{r}
setVisualStyle('Marquee')
```

### My Own Style
Create your own style with node attribute fill mappings and some defaults
```{r}
style.name = "myStyle"
defaults <- list(NODE_SHAPE="diamond",
                 NODE_SIZE=30,
                 EDGE_TRANSPARENCY=120,
                 NODE_LABEL_POSITION="W,E,c,0.00,0.00")
nodeLabels <- mapVisualProperty('node label','id','p')
nodeFills <- mapVisualProperty('node fill color','group','d',c("A","B"), c("#FF9900","#66AAAA"))
arrowShapes <- mapVisualProperty('Edge Target Arrow Shape','interaction','d',c("activates","inhibits","interacts"),c("Arrow","T","None"))
edgeWidth <- mapVisualProperty('edge width','weight','p')

createVisualStyle(style.name, defaults, list(nodeLabels,nodeFills,arrowShapes,edgeWidth))
setVisualStyle(style.name)
```

### Browse Available Functions, Commands and Arguments
RCy3 functions
```{r}
help(package=RCy3)
```

Open swagger docs for live instances of CyREST API and Commands API:
```{r}
cyrestAPI()  # CyREST API
commandsAPI()  # Commands API
```

List available commands and arguments in R. Use "help" to list top level:
```{r}
commandsHelp("help")  
```

List **network** commands. Note that "help" is optional:
```{r}
commandsHelp("help network")  
```

List arguments for the **network select** command:
```{r}
commandsHelp("help network select") 
```

Those are the basics. You are now ready to work with some real data! See other basic and advanced R tutorials, https://github.com/cytoscape/cytoscape-automation/tree/master/for-scripters/R.
